Active Learning from Blackbox to Timed Connectors

### 2016-01-18 paper/paper.tex
.

@@

\begin{itemize}
  \item [-] A \emph{Sync} channel accepts a data item from its source end iff the data item can be
    dispensed through its sink end simultaneously.
  \item [-] A \emph{Lo ssySync} channel is always ready to accept data items. These items will be
    dispensed through its sink end if possible, otherwise they will be lost.
  \item [-] A \emph{SyncDrain} channel has two source ends and no sink end. It can accept a data
    item through one of its source end iff a data item is also available to be accepted
    simultaneously through the other end. Then both items will be dropped.
  \item [-] A \emph{FIFO1} channel is an asychronous channel with a buffer cell. It accepts a
    data item whenever the buffer is empty. 
  \item [-] A t-timer channel accepts any data item from its source end, and later dispense it to
    its sink end after a delay of $t$ time units.

@@

A Mealy machine is in some state $q\in S$ is always ready for inputs at any time. If an input symbol
$i\in I$ is provided, the Mealy machine generates output $o=\lambda(q,i)$ and jumps to state
$q'=\delta(q,i)$. Such an edge is also denoted as $q\xrightarrow[]{i/o}q'$.

@@

\begin{definition}[Production]
  Now we're going to define the production operator \emph{prod} of two parameterized Mealy machines
  \[
  prod(PM_1,PM_2)=PM_3
  \]
  as follows. Here we assume that $PM_2.O\cap PM_1.I=\varnothing$
  \begin{itemize}
  	\item[-] $\forall\Sigma, PM_3.S(\Sigma)=PM_1.S(\Sigma)\times PM_2.S(\Sigma)$
    \item[-] $PM_3.I=PM_1.I\cup PM_2.I-PM_1.O$
    \item[-] $PM_3.O=PM_1.O\cup PM_2.O-PM_2.I$
  \end{itemize}
  \emph{(Here we assume that sink ends of $PM_1$ can be connected to source ends $PM_2$, but not
  vise versa)}
  \begin{itemize}
    \item[-] $PM_3.s_0=(PM_1.s_0, PM_2.s_0)$
    \item[-] $\forall\Sigma, PM_3.\delta(\Sigma)((s_1,s_2), i)=$
      \begin{displaymath}
        \left\{
        \begin{array}[ht]{ll}
          (Out_1 + Out_2)\downarrow_{PM_3.O} & Out_1\neq\rblock\land Out_2\neq\rblock \\
          \rempty & i = T \\
          \rblock & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
      where we have
      \begin{itemize}
        \item[*] $In_1 = i\downarrow_{PM_1.I}$
        \item[*] $Out_1 = PM_1.\delta(\Sigma)(s_1,In_1)$
        \item[*] $In_2 = (Out_1 + i)\downarrow_{PM_2.I}$
        \item[*] $Out_2 = PM_2.\delta(\Sigma)(s_2,In_2)$
      \end{itemize}
    \item[-] $\forall\Sigma, PM_3. \lambda(\Sigma)((s_1,s_2),i)=(s_1',s_2')$
      where we have
      \begin{itemize}
        \item[*] $s_1' = PM_1.\lambda(\Sigma)(s_1,In_1)$
        \item[*] $s_2' = PM_2.\lambda(\Sigma)(s_2,In_2)$
      \end{itemize}
  \end{itemize}
\end{definition}

The idea in \emph{production} is quite simple. The output of $PM_1$ would be provided as part of the
input of $PM_2$. Time actions T is only executed simutaneously between $PM_1$ and $PM_2$.

As mentioned in the notation above, we cannot connect a sink end of $PM_2$ to a source end of
$PM_1$, which makes it difficult to construct connectors like alternator in Figure
\ref{fig:reoconnector}. To address the problem, we define a \emph{link} operation to connect sink
ends and source ends in the same connector.

\begin{definition}[Link Operator]
  The \emph{link} operation constructs connector by connecting a sink end $OUT$ to a source end $IN$
  within the same connector.   
  \[
  PM' = link(PM, OUT, IN)
  \]
  Here we asuume that $IN\in PM.I$ and $OUT\in PM.O$.

  \begin{itemize}
  	\item[-] $\forall\Sigma, PM'.S(\Sigma)=PM.S(\Sigma)$
    \item[-] $PM'.I=PM.I-\{IN\}$
    \item[-] $PM'.O=PM.O-\{OUT\}$
    \item[-] $PM'.s=PM.s$
    \item[-] $\forall\Sigma, i\in Input(\Sigma,PM'.I,PM'.O)$, $PM'.\delta(\Sigma)(s, i)=$
      \begin{equation}
        \label{equ:existd}
        PM.\delta(\Sigma)(s,T)\downarrow_{PM'.O}
      \end{equation}
      if $i=T$. When $i$ is a non-time action, we need to check whether there's a data item
      $d\in\Sigma$
      satisfies
      \[
      PM.\delta(\Sigma)(s,i\cup\{IN:d\})\downarrow_{OUT} = d
      \]
      if the formula is satisfied, we have $PM'.\delta(\Sigma)(s, i)=$
      \[
      PM.\delta(\Sigma)(s,i\cup\{IN:d\})\downarrow_{PM'.O}
      \]
      otherwise $PM'.\delta(\Sigma)(s, i)=\rblock$
    \item[-] $\forall\Sigma, i\in Input(\Sigma,PM'.I,PM'.O).'$ the transition
      function can be defined as $PM.\lambda(\Sigma)(s,i)=$
      \begin{displaymath}
        \left\{
        \begin{array}[ht]{ll}
          PM.\lambda(\Sigma,T) & i=T \\
          PM.\lambda(\Sigma,i\cup\{IN:d\}) & \exists d\in\Sigma \mbox{ satisfes Equation
          \ref{equ:existd}} \\
          s & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
  \end{itemize}
\end{definition}

With \emph{link} and \emph{prop} defined, complex connectors can be formed by simple channels in
form of \emph{parameterized Mealy machines}. It will be transformed into a concrete Mealy machine
later once the alphabet $\Sigma$ is provided.

>>>>

\newcommand{\OUT}[0]{\mbox{OUT}}
\newcommand{\IN}[0]{\mbox{IN}}

@@

\begin{itemize}
  \item [-] A \emph{Sync} channel accepts a data item from its source end iff the data item can be
    dispensed through its sink end simultaneously.
  \item [-] A \emph{LossySync} channel is always ready to accept data items. These items will be
    dispensed through its sink end if possible, otherwise they will be lost.
  \item [-] A \emph{SyncDrain} channel has two source ends and no sink end. It can accept a data
    item through one of its source end iff a data item is also available to be accepted
    simultaneously through the other end. Then both items will be dropped.
  \item [-] A \emph{FIFO1} channel is an asychronous channel with a buffer cell. It accepts a
    data item whenever the buffer is empty. 
  \item [-] A t-timer channel accepts any data item from its source end, and later dispense it to
    its sink end after a delay of $t$ time units.

@@

@@

\begin{definition}[Production]
  The production operator \emph{prod} of two parameterized Mealy machines
  \[
  \mathcal{PM'} = prod(\mathcal{PM}_1,\mathcal{PM}_2)\mbox{ if }\mathcal{PM}_2.O\cap
  \mathcal{PM}_1.I=\varnothing
  \]
  is defined as follows.
  \begin{itemize}
  	\item[-] $\forall\Sigma. \mathcal{PM}'.S(\Sigma)=\mathcal{PM}_1.S(\Sigma)\times \mathcal{PM}_2.S(\Sigma)$
    \item[-] $\mathcal{PM}'.I=(\mathcal{PM}_1.I\cup \mathcal{PM}_2.I)\backslash \mathcal{PM}_1.O$
    \item[-] $\mathcal{PM}'.O=(\mathcal{PM}_1.O\cup \mathcal{PM}_2.O)\backslash \mathcal{PM}_2.I$
  \end{itemize}
  \emph{(Here we assume that sink ends of $\mathcal{PM}_1$ can be connected to source ends $\mathcal{PM}_2$, but not
  vise versa)}
  \begin{itemize}
    \item[-] $\mathcal{PM}'.s_0=(\mathcal{PM}_1.s_0, \mathcal{PM}_2.s_0)$
    \item[-] $\forall\Sigma. \mathcal{PM}'.\delta(\Sigma)((s_1,s_2), i)=$
      \begin{displaymath}
        \left\{
        \begin{array}[ht]{ll}
          (Out_1 \cup Out_2)\downarrow_{\mathcal{PM}'.O} & Out_1\neq\rblock\land Out_2\neq\rblock \\
          \rempty & i = T \\
          \rblock & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
      where we have
      \begin{itemize}
        \item[*] $In_1 = i\downarrow_{\mathcal{PM}_1.I}$
        \item[*] $Out_1 = \mathcal{PM}_1.\delta(\Sigma)(s_1,In_1)$
        \item[*] $In_2 = (Out_1 \cup i)\downarrow_{\mathcal{PM}_2.I}$
        \item[*] $Out_2 = \mathcal{PM}_2.\delta(\Sigma)(s_2,In_2)$
      \end{itemize}
    \item[-] $\forall\Sigma. \mathcal{PM}'. \lambda(\Sigma)((s_1,s_2),i)=(s_1',s_2')$
      where we have
      \begin{itemize}
        \item[*] $s_1' = \mathcal{PM}_1.\lambda(\Sigma)(s_1,In_1)$
        \item[*] $s_2' = \mathcal{PM}_2.\lambda(\Sigma)(s_2,In_2)$
      \end{itemize}
  \end{itemize}
\end{definition}

The idea in \emph{production} is quite simple. The output of $\mathcal{PM}_1$ would be provided as part of the
input of $\mathcal{PM}_2$. Time actions T is only executed simutaneously between $\mathcal{PM}_1$ and $\mathcal{PM}_2$.

As mentioned in the notation above, we cannot connect a sink end of $\mathcal{PM}_2$ to a source end of
$\mathcal{PM}_1$, which makes it difficult to construct connectors like alternator in Figure
\ref{fig:reoconnector}. To address the problem, we define a \emph{link} operation to connect sink
ends and source ends in the same connector.

\begin{definition}[Link]
  The \emph{link} operator constructs connector by connecting a sink end $\OUT$ to a source end $\IN$
  within the same connector.   
  \[
  \mathcal{PM}' = link(\mathcal{PM}, \OUT, \IN)
  \]
  Here we assume that $\IN\in \mathcal{PM}.I$ and $\OUT\in \mathcal{PM}.O$.

  \begin{itemize}
  	\item[-] $\forall\Sigma. \mathcal{PM}'.S(\Sigma)=\mathcal{PM}.S(\Sigma)$
    \item[-] $\mathcal{PM}'.I=\mathcal{PM}.I\backslash\{\IN\}$
    \item[-] $\mathcal{PM}'.O=\mathcal{PM}.O\backslash\{\OUT\}$
    \item[-] $\mathcal{PM}'.s_0=\mathcal{PM}.s_0$
    \item[-] $\forall\Sigma, i\in
      Input(\Sigma,\mathcal{PM}'.I,\mathcal{PM}'.O).\mathcal{PM}'.\delta(\Sigma)(s,i)=$
      \begin{displaymath}
        \small
        \left\{
        \begin{array}[h]{lr}
          \rempty & i=T \\
          \mathcal{PM}.\delta(\Sigma)(s,i\cup\{\IN:d\})\downarrow_{\mathcal{PM}'.O} & (*)\\
          \rblock & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
      where the condition (*) is defined as
      \[
      \exists d.\mathcal{PM}.\delta(\Sigma)(s,i\cup\{\IN:d\})(\OUT)=d
      \]
    \item[-] $\forall\Sigma, i\in
      Input(\Sigma,\mathcal{PM}'.I,\mathcal{PM}'.O).$
      \begin{displaymath}
        \small
        \mathcal{PM}.\lambda(\Sigma)(s,i)=
        \left\{
        \begin{array}[ht]{lr}
          \mathcal{PM}.\lambda(\Sigma,T) & i=T \\
          \mathcal{PM}.\lambda(\Sigma,i\cup\{\IN:d\}) & (*)\\
          s & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
  \end{itemize}
\end{definition}

With \emph{link} and \emph{prop} defined, a connector can be formed by simpler ones in
form of \emph{parameterized Mealy machines}. It will be transformed into a concrete Mealy machine
later, once the alphabet $\Sigma$ is provided.
